Evolution of ORM Systems: Formal Foundations

### 2013-01-16 doc/description.tex
rewrited decomposition of SW

\section{Transformation for Models Manipulation}
The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppT$ of transformations) and database (set $DbT$). These sets are defined in this section.
\subsection{Application Transformation}
This section provide transformations for manipulating the application layer of a software - the $AppT$ set. The set contains not only transformations for model creation, alternation and decomposition, but some operations to obtain information about a model too.

@@

@@

@@

\subsection{Software Decomposition}
\begin{align*}
\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
\Phi RemoveClass(l, s) = database(s) \bullet dropClass(l) \\
\\ l \in Label, s \in Software 
\end{align*}
\begin{align*}
dropClass(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
\end{align*}

\begin{align*}
\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
\Phi RemoveProperty(l, l_p, s) =  database(s) \bullet dropProp(l, l_p) \end{align*}
\begin{align*}
dropProp(p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
} \\ \\
  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
 \end{cases}
\end{align*}

\begin{align*}
\mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
\Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
\Phi RemoveAssociation(l, l_a, s) = database(s) \bullet dropAssoc(l, l_a) \\
l_c, l_a \in Labels, a \lhd Applications
\end{align*}
\begin{align*}
dropAssoc(l, l_a, d) =  \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a
 \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
 \\ \\
 \inference{\exists t \in tables(d) : name(t) = l_a
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l
 \\ \wedge \forall r \in data(d) : reference(r) \neq t}{ tables(d) \setminus t}
 \end{cases}
\end{align*}

\subsection{Software Alternation}
\paragraph{Rename Class}
\begin{align*}
\mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterClass(l_o, l_n) \\ \\
\end{align*}
\begin{align*}
alterTable:  Label \times Label \times Database \rightarrow Database \\
alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
\end{gathered}}\\
l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
\end{align*}

\paragraph{Rename Property}
\begin{align*}
	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
\Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{alterTable(l_o, l_n, database(s))}
 \end{cases}\\
l_c, l_o, l_n \in Label,  s \lhd Software
\end{align*}

>>>>

\section{Transformation for Models Manipulation}
The evolution affects both parts of software and as shown in section \ref{sec:evoIntro} it is mapped to transformation specific for application (set $AppT$ of transformations) and database (set $DbT$). These sets are defined in this section.

\subsection{Application Transformation}
This section provide transformations for manipulating the application layer of a software - the $AppT$ set. The set contains not only transformations for model creation, alternation and decomposition, but some operations to obtain information about a model too.

@@

\paragraph{Alter Table}
\begin{align*}
alterTable:  Label \times Label \times Database \rightarrow Database \\
alterTable(l_o, l_n, d) = \inference{\exists t \in tables(d) : name(d) = l_o \\ \wedge \forall e \in tables(d) : name(e) \neq l_n}{\begin{gathered}
t.name = l_n \\ \wedge \forall e \in tables(a) \wedge \forall (f \in fks(e) \wedge reference(f) = l_o ) \\ : f.reference = l_n 
\end{gathered}}\\
l_o, l_n \in Label, s \in States, a \lhd Applications, d \lhd Database
\end{align*}

@@

\paragraph{Remove table}
\begin{align*}
	dropTable(l,d) = \inference{ \exists t \in tables(d) : name(t) = l  \\ \wedge \not \exists r \in data(d) : owningTable(r) = l \\  \not \exists j \in tables(d) \wedge f \in foreignKeys(j) : reference(f) = l }{ tables(d) \setminus \rho(c)}
\end{align*}

\paragraph{Remove column}
\begin{align*}
	dropColumn(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in columns(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{columns(t) \setminus c}\\
	 l \in Label, d \lhd Database 
\end{align*}

\paragraph{Remove Foreign Key}
\begin{align*}
	dropKey(l, l_p, d) = \inference{ \exists t \in tables(d) : name(t) = l \\ \wedge \exists c \in foreignKeys(t) : name(c) = l_p \\ \wedge \forall e \in  pairs(r), r \in tableData(t, d) : \\ reference(e) \neq l_p \vee value(e) = \varnothing  }{foreignKeys(t) \setminus c}\\
	 l \in Label, d \lhd Database 
\end{align*}

@@

\subsection{Software Decomposition}
\begin{align*}
	\mathbf{RemoveClass}: Label \times Software \rightarrow State  \\
	\Psi RemoveClass(l, s) = applications(s) \circ remClass(l) \\
	\Phi RemoveClass(l, s) = database(s) \bullet dropTable(l) \\
	l \in Label, s \in Software 
\end{align*}

\begin{align*}
	\mathbf{RemoveProperty} : Label \times Label \times State \rightarrow State \\
	\Psi RemoveProperty(l, l_p, s) =  applications(s) \circ  remProp(l, l_p) \\
	\Phi RemoveProperty(l, l_p, s) = \begin{cases}
		 \inference{cardinality(propertyByName(l_p, application(s)) \leq 1}{database(s) \bullet dropColumn(l, l_p)}  \\\\
		 \inference{cardinality(propertyByName(l_p, application(s)) >1}{database(s) \bullet dropTable(l)}
 	\end{cases}\\
 	l, l_p \in Label, s \lhd Software
\end{align*}

\begin{align*}
    \mathbf{RemoveAssociation} : Label \times Labels \times State \rightarrow State \\ 
    \Psi RemoveAssociation(l, l_a, s) = application(s) \circ remAss(l, l_a) \\
    \Phi RemoveAssociation(l, l_a, s) = \begin{cases}
     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) \leq 1 \\}{database(s) \bullet dropForeignKey(l_a, reference(a))} \\\\
     	\inference{a = associationByName(l, l_a, application(s)) \\ \wedge startCardinality(a) > 1}{database(s) \bullet dropTable(l_a)} 
     \end{cases}\\
    l_c, l_a \in Labels, s \lhd Software
\end{align*}

\subsection{Software Alternation}
\paragraph{Rename Class}
\begin{align*}
\mathbf{RenameClass} : Label \times Label \times State \rightarrow State \\ 
\Psi(RenameClass(l_o, l_n), s) = application(s) \circ renClass(l_o, l_n) \\
\Phi(RenameClass(l_o, l_n), s) = database(s) \bullet alterTable(l_o, l_n) \\ \\
\end{align*}

\paragraph{Rename Property}
\begin{align*}
	\mathbf{RenameProperty} : Label \times Label \times Label \times State \rightarrow State \\ \\
\Psi(RenameProperty(l_c, l_o, l_n, s)) = application(s) \circ renProp(l_c, l_o, l_n) \\
\Phi(RenameProperty(l_c, l_o, l_n, s)) = \begin{cases}
 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) \leq 1}{database(s) \bullet alterColumn(l_c, l_o, l_n))} \\\\
 	\inference{cardinality(propertyByName(l_c, l_o, application(s))) > 1}{database(s) \bullet alterTable(l_o, l_n, database(s))}
 \end{cases}\\
l_c, l_o, l_n \in Label,  s \lhd Software
\end{align*}
